\appendix
\section{Review protocol details}
\label{Appendix}


\subsection{PICOC and Search String}
\label{subsec:Appendix-PICOC-and-Search-String}
\textcolor{orange}{Although the search strings reported below are not identical in their textual form, they implement the same conceptual search strategy across all selected digital libraries. The differences are due to database-specific query syntax, field naming conventions, and interface limitations, which required adapting the formulation of the search string to each search engine while preserving the intended scope of the review as closely as possible.}
\input{tables/PICOC}
\input{tables/search-string}


\subsection{Inclusion and Exclusion Criteria}
\label{subsec:Appendix-IC-EC}
The EC/IC criteria of our study are as follows.

\textit{Exclusion Criteria}.
%Publications are excluded if any of the following EC apply:
\begin{enumerate}[label=(EC\arabic*), leftmargin=19.5mm, itemsep=1pt, parsep=0pt]
    \item The study is a position paper or a short paper.
    \item The study is not peer-reviewed.
    \item The study is a duplicate or an earlier version of another candidate paper.
    \item The full text is not available.
    \item The study is a secondary or tertiary study (Review, mini-review, survey, etc.) 
    \item The study tests other authors' methodology (e.g., a study that tries to replicate the results of another study)
\end{enumerate}

\textit{Inclusion Criteria}. 
%In order to be included, a publication has to satisfy all the following IC:
\begin{enumerate}[label=(IC\arabic*), leftmargin=19.5mm, itemsep=1pt, parsep=0pt]
\item The study is written in English.
\item The study focuses on Customer Churn Prediction in the Retail Sector using AI, based on customer and purchase data, which can be detailed as follows:
\begin{enumerate}[label=(IC2.\arabic*), leftmargin=19.5mm, itemsep=1pt, parsep=0pt]
\item The study is strictly about Customer Churn prediction.
\item The domain is strictly about retail.
\item The paper strictly uses AI methodologies for churn prediction.
\item Customer churn prediction is based on data about customers’ purchases (transactions) AND/OR customer profiling.
\end{enumerate}
\end{enumerate}